---
title: "BoneCellType: R package for bone-relevant cell type assignment"
author: Kristyna Kupkova
date: "`r Sys.Date()`"
#output: rmarkdown::html_vignette
output: 
  prettydoc::html_pretty:
    theme: hpstr
    toc: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{BoneCellType: R package for bone-relevant cell type assignment}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r klippy, echo=FALSE, include=TRUE}
# add copy clipboard button
klippy::klippy()
```

```{r, echo=FALSE}
# These settings make the vignette prettier
knitr::opts_chunk$set(results="hold", collapse=FALSE, message=FALSE, 
                      warning = FALSE)
#refreshPackage("GenomicDistributions")
#devtools::build_vignettes("code/GenomicDistributions")
#devtools::test("code/GenomicDistributions")
```

# Introduction to BoneCellType

The BoneCellType package aims to help assign cell types to single-cell
RNA-seq clusters derived from bone samples. BoneCellType uses a
custom-assembled database of bone-relevant markers (assembly approach
described [here](https://github.com/kkupkova/Mouse-bone-markers)). There
are two approaches introduced here to reach this goal: 1) marker
overrepresentation analysis using Fisher's exact test, and 2) gene
(marker) set enrichment analysis.

**Note!** The results produced with BoneCellType do not assign
definitive cell-type labels; only an initial prediction based on
statistical testing and further manual curation is needed.

# Setup

Let's load the BoneCellType package along with a few other packages
whose functions are used in this tutorial.

The input into the functions presented here is a data.frame with
variable markers (i.e. features) for individual clusters that can be
obtained by using [Seurat](https://satijalab.org/seurat/) `FindMarkers`
function. Here we provide a previously generated example dataset.

```{r setup}
library(BoneCellType)
library(dplyr)
library(ggplot2)

# example dataset
varMarkerFile = system.file("extdata", "cluster_variable_features.csv", 
                            package = "BoneCellType")
varMarkers = read.csv(varMarkerFile, sep = ",")

# show top rows of the example dataset
head(varMarkers)
```

# 1. Marker overrepresentation analysis

To calculate marker overrepresentation use `markerEnrich` function. It
automatically provides a database of bone-relevant markers. Results can
be further dirrectly plotted with an associated `plotMarkerEnrich`
function, output of which is an editable *ggplot* object.

**Note**: While the default setting of `markerEnrich` use all available
variable markers per cluster, based on practical experience, more
reliable results can be achieved by selecting only few top markers by
setting the parameter *topN* (as shown in the example below). Default
mode selects the top markers based on log2FC, this can be changed by
setting the parameter *sortBy="padj* ". Changing the *sortBy*, however,
usually does not change much.

```{r marker-overrep-top10}
# run overrepresentation analysis uasing top 10 variable markers (features) per cluster
ORres = markerEnrich(varMarkers, topN = 10)

# plot the top 5 most significant results
plotMarkerEnrich(ORres, topN = 5)
```

The labels on the x-axis contain following information \<cluster number
\_\_ source database \_\_ cell type\>. Dashed line indicates padj =
0.05.

Let's now have a look at the format of the `markerEnrich` results.

```{r topRes}
# show top 2 significant result for each cluster
topORres = ORres %>% 
  group_by(cluster) %>% 
  top_n(n = 2, wt = neg_log10_padj)

topORres
```

As shown, the output is a table with following variables- *cluster*:
cluster number / *cellType*: cell type / *database*: source database
from which markers originate / *p*: p-value / *oddsRatio*: odds ratio /
*n_olaps*: number of overlap between variable markers (features) and the
marker database / *overlaps*: list of individual overlapping markers /
*padj*: adjusted p-value /*neg_log10_padj*: -log10(padj).

**Recommendation:** Try few different settings of the `markerEnrich`
function try different *topN*, *sortBy* arguments. In general I wouldn't
set the *byDatabase* to FALSE, as the TISSUES database has relatively a
lot of markers (some of which I am not too sure about) and it can skew
the results. But overall I definitely recommend to explore the whole
result table, to see which markers drive associated results and if that
cell type assignment then makes sense.

# 2. Gene (marker) set enrichment analysis
